---
title: "Stratified PH using saswine inside a Rstudio notebook"
output: html_notebook
---

The data comes from "The Effect of 6-Mercaptopurine on the Duration of Steroid-induced Remissions in Acute Leukemia: A Model for Evaluation of Other Potentially Useful Therapy" (Freireich and al, Blood 1963 21:699-716)

21 ordered times to relapse  (in weeks) are compared among children in remission from acute leukemia. Matching was on remission status and hospital (pairid): two patients come from the same hospital, with one treated and the other untreated.

Goal: Estimate the relative risk for 6-MP versus placebo, adjusting for hospital and remission status

# Data input and cleaning

```{r engine='sas', engine.path='/usr/local/bin/saswine', result='asis'}
data ref.leukemia;
input ID$ pairid   tpbo   t6mp   death; 
datalines;
"1" 1 1 10 1
"2" 2 22 7 1
"3" 3 3 32 0
"4" 4 12 23 1
"5" 5 8 22 1
"6" 6 17 6 1
"7" 7 2 16 1
"8" 8 11 34 0
"9" 9 8 32 0
"10" 10 12 25 0
"11" 11 2 11 0
"12" 12 5 20 0
"13" 13 4 19 0
"14" 14 15 6 1
"15" 15 8 17 0
"16" 16 23 35 0
"17" 17 5 6 1
"18" 18 11 13 1
"19" 19 4 9 0
"20" 20 1 6 0
"21" 21 8 10 0
;
PROC PRINT DATA=ref.leukemia; 
   TITLE 'Leukemia input'; 
RUN; 

```

```{r engine='sas', engine.path='/usr/local/bin/saswine', result='asis'}
PROC TRANSPOSE DATA=ref.Leukemia OUT=ref.LEU1;
BY PAIRID;
VAR tpbo;
RUN;

DATA ref.LEUtpbo;
SET ref.LEU1 (rename=(COL1=time) drop=_name_) ;
death=1; 
group=2;
RUN;

PROC PRINT DATA=ref.LEUtpbo;
   TITLE 'Survival of untreated Leukemia patients';
RUN;
```

```{r engine='sas', engine.path='/usr/local/bin/saswine', result='asis'}
PROC TRANSPOSE DATA=ref.Leukemia OUT=ref.LEU2;
BY PAIRID;
VAR  T6MP ;
RUN;
PROC TRANSPOSE DATA=ref.Leukemia OUT=ref.LEU3;
BY PAIRID;
VAR death;
RUN;

DATA ref.LEUt6mp; 
MERGE ref.LEU2 (rename=(COL1=time) drop=_name_) ref.LEU3 (rename=(COL1=death) drop=_name_); 
by PAIRID;
group=1;
RUN;

PROC PRINT DATA=ref.LEUt6mp;
TITLE 'Survival of treated Leukemia patients';
RUN;
```

```{r engine='sas', engine.path='/usr/local/bin/saswine', result='asis'}
DATA ref.Leukemia_final;
SET ref.LEUtpbo ref.LEUt6mp; 
by PAIRID; 
RUN;

PROC PRINT DATA=ref.Leukemia_final;
title 'Leukemia patients grouped';
run;
```

# Correct analysis

```{r engine='sas', engine.path='/usr/local/bin/saswine', result='asis'}
PROC PHREG DATA=ref.Leukemia_final;
MODEL TIME*DEATH(0)=  GROUP /RL;
STRATA PAIRID;
RUN;
```

The risk of relapse is 6 times higher in the placebo group that in the 6MP treated group
```{r}
exp(1.79176)
```

# Incorrect analysis:

```{r engine='sas', engine.path='/usr/local/bin/saswine', result='asis'}
PROC PHREG DATA=ref.Leukemia_final;
MODEL TIME*DEATH(0)=  GROUP /RL; 
RUN;
```

The analysis is incorrect because it is not clear what the reference population is.

```{r engine='sas', engine.path='/usr/local/bin/saswinehtmls', result='asis'}
proc lifetest data=ref.Leukemia_final plots=(s,h) METHOD=KM NOPRINT NOTABLE;
time time*death(0);
strata group;
title'Estimating Survival Time';
run;
```

```{r echo=FALSE, message=FALSE}
shiny::includeHTML('~/.wine-sas91/drive_c/SASWINE/chunk.html')
```


